---
title: "Stats 2 - Project 2"
author: "Jon Paugh"
date: "3/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data Source:
https://archive.ics.uci.edu/ml/datasets/Bank%20Marketing

Original research on this dataset, and source:
Moro et al., 2014] S. Moro, P. Cortez and P. Rita. A Data-Driven Approach to Predict the Success of Bank Telemarketing. Decision Support Systems, Elsevier, 62:22-31, June 2014

```{r}

#install.packages("ROCR")

```


# Load the data



```{r}

BankData <- read.csv(file.choose(), sep=";")

BankData


```

# EDA
```{r}

summary(BankData)

sapply(BankData, class)

```




## Correlation Plot

```{r}
suppressMessages(library(corrplot))


numVars <- sapply(BankData, is.numeric)
correlationMatrix <- cor(BankData[,numVars])
# same thing but now with p value

corrplot(correlationMatrix,main='\nCorrelation Plot with P Values', type="upper", 
         sig.level = 0.05, tl.cex = 0.75, number.cex=1)


numVars


```

# check the data


## numeric data


```{r}

library("ggplot2")
library("ggthemes")

NiceBoxPlot <- function(columnBy, columnToShow, columnByDesc, columnToShowDesc)
{


ggplot(BankData, aes(x=BankData[,columnBy], y=BankData[,columnToShow])) + geom_boxplot() +
    stat_summary(fun.y=mean, geom="point", shape=5, size=4) +
  ggtitle(paste(columnByDesc, "by",columnToShowDesc)) + xlab("Outcome") + ylab(columnToShowDesc) +
  theme_economist()

}

NiceBoxPlot("y", "duration", "Outcome", "duration - Duration of Last Contact (secs)")
NiceBoxPlot("y", "campaign", "Outcome", "campaign - Number of Contacts During this Campaign")
NiceBoxPlot("y", "pdays", "Outcome", "pdays - Number of Days Since Previous Contact")
NiceBoxPlot("y", "previous", "Outcome", "previous - Previous Number of Contacts")
NiceBoxPlot("y", "emp.var.rate", "Outcome", "emp.var.rate - Employment Variation Rate (quaterly indicator)")
NiceBoxPlot("y", "cons.price.idx", "Outcome", "cons.price.idx - Consumer Price Index (montly indicator)")
NiceBoxPlot("y", "cons.conf.idx", "Outcome", "cons.conf.idx - Consumer Confidence Index(monthly indicator)")
NiceBoxPlot("y", "euribor3m", "Outcome", "euribor3m - eurobor 3 month rate")
NiceBoxPlot("y", "nr.employed", "Outcome", "nr.employed - number of employees (quarterly indicator)")


```




```{r}

library(ggplot2)
library(ggthemes)
ggplot(data=BankData, aes(BankData$y)) + 
  geom_bar(aes(y = (..count..)/sum(..count..))) + ylab("percent") + theme_economist() + ggtitle("Histogram of Outcome") + xlab("Outcome") 
ggplot(data=BankData, aes(BankData$job, fill=BankData$y)) + 
  geom_bar(aes(y = (..count..)/sum(..count..)), position="dodge") + ylab("percent") + theme_economist() + ggtitle("Histogram of Jobs") + xlab("Job") + scale_fill_discrete(name = "Output:")

ggplot(data=BankData, aes(BankData$marital, fill=BankData$y)) + 
  geom_bar(aes(y = (..count..)/sum(..count..)), position="dodge") + ylab("percent") + theme_economist() + ggtitle("Histogram of Marital Status") + xlab("Marital Status") + scale_fill_discrete(name = "Output:")





```



## use fuction to show percentages for categorical data

```{r}

library("dplyr")

ShowPieChart <- function(columnBy, columnToShow, titleName)
{
  df <- dplyr::group_by(BankData, .dots = c(columnBy, columnToShow)) %>%
    dplyr::summarize(counts = n()) %>%
    dplyr::mutate(perc = (counts / sum(counts)) * 100) %>%
    dplyr::arrange_(.dots=c(columnBy, columnToShow))
  
  
  # preparing the plot
  ggplot2::ggplot(df, aes('', counts)) +
    geom_col(
      position = 'fill',
      color = 'black',
      width = 1,
      aes(fill = y)
    ) +
    ggtitle(titleName) +
    facet_wrap(paste("~",columnBy), labeller = "label_both") +
    geom_label(
      aes(label = paste0(round(perc), "%"), group = "y"),
      position = position_fill(vjust = 0.5),
      color = 'black',
      size = 5,
      show.legend = FALSE
    ) + scale_fill_discrete(name = "Outcome:") +
    coord_polar(theta = "y")
}

ShowPieChart("job", "y", "Outcome by Job")
ShowPieChart("marital", "y", "Outcome by Marital Status")
ShowPieChart("education", "y", "Outcome by Education")
ShowPieChart("housing", "y", "Outcome by Housing")
ShowPieChart("default", "y", "Outcome by Credit In Default")
ShowPieChart("loan", "y", "Outcome by loan status")
ShowPieChart("contact", "y", "Outcome by Contact")
ShowPieChart("poutcome", "y", "Outcome by Day of the Week")


```




## Check of the range of some of these continuous variables with histograms

```{r}

ggplot(data=BankData,aes(x=BankData$duration)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$duration, na.rm = TRUE), sd = sd(BankData$duration, na.rm = TRUE))) +
  labs(x = "duration", y = "Count", title = "Distribution of duration") +   theme_economist()

ggplot(data=BankData,aes(x=BankData$campaign)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$campaign, na.rm = TRUE), sd = sd(BankData$campaign, na.rm = TRUE))) +
  labs(x = "campaign", y = "Count", title = "Distribution of campaign") +   theme_economist()

ggplot(data=BankData,aes(x=BankData$pdays)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$campaign, na.rm = TRUE), sd = sd(BankData$campaign, na.rm = TRUE))) +
  labs(x = "pdays", y = "Count", title = "Distribution of pdays") +   theme_economist()


ggplot(data=BankData,aes(x=BankData$previous)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$previous, na.rm = TRUE), sd = sd(BankData$campaign, na.rm = TRUE))) +
  labs(x = "previous", y = "Count", title = "Distribution of previous") +   theme_economist()


ggplot(data=BankData,aes(x=BankData$emp.var.rate)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$emp.var.rate, na.rm = TRUE), sd = sd(BankData$emp.var.rate, na.rm = TRUE))) +
  labs(x = "emp.var.rate", y = "Count", title = "Distribution of emp.var.rate") +   theme_economist()

ggplot(data=BankData,aes(x=BankData$cons.price.idx)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$emp.var.rate, na.rm = TRUE), sd = sd(BankData$cons.price.idx, na.rm = TRUE))) +
  labs(x = "cons.price.idx", y = "Count", title = "Distribution of cons.price.idx") +   theme_economist()

ggplot(data=BankData,aes(x=BankData$cons.conf.idx)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$cons.conf.idx, na.rm = TRUE), sd = sd(BankData$cons.conf.idx, na.rm = TRUE))) +
  labs(x = "cons.conf.idx", y = "Count", title = "Distribution of cons.conf.idx") +   theme_economist()

ggplot(data=BankData,aes(x=BankData$euribor3m)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$euribor3m, na.rm = TRUE), sd = sd(BankData$euribor3m, na.rm = TRUE))) +
  labs(x = "euribor3m", y = "Count", title = "Distribution of euribor3m") +   theme_economist()

ggplot(data=BankData,aes(x=BankData$nr.employed)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$nr.employed, na.rm = TRUE), sd = sd(BankData$nr.employed, na.rm = TRUE))) +
  labs(x = "nr.employed", y = "Count", title = "Distribution of nr.employed") +   theme_economist()


```
### comments on results
Duration looks normal but right skewed
Campaign is NOT normal - it looks to be decreasing log scale
pdays is not really a numeric - the values are only 0 or 1000 - we can condense to buckets
previous also only have a hanful of values, can probably condense
emp.var.rate - might be able to reduce to buckets as well? It only have a couple values
cons.price.idx - nice distribution
cons.conf.idx - nice distribution
euribor3m - maybe buckets as well. There's some distribution but not much




### It would be good to show the same values with separate histograms for the two y/Response values

```{r}

ggplot(data=BankData,aes(x=BankData$duration, fill=BankData$y)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$duration, na.rm = TRUE), sd = sd(BankData$duration, na.rm = TRUE))) +
  labs(x = "duration", y = "Count", title = "Distribution of duration") +   theme_economist() + geom_density(alpha=.3) + scale_fill_discrete(name = "Attrition:")

ggplot(data=BankData,aes(x=BankData$campaign, fill=BankData$y)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$campaign, na.rm = TRUE), sd = sd(BankData$campaign, na.rm = TRUE))) +
  labs(x = "campaign", y = "Count", title = "Distribution of campaign") +   theme_economist() + geom_density(alpha=.3) + scale_fill_discrete(name = "Attrition:")

ggplot(data=BankData,aes(x=BankData$pdays, fill=BankData$y)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$campaign, na.rm = TRUE), sd = sd(BankData$campaign, na.rm = TRUE))) +
  labs(x = "pdays", y = "Count", title = "Distribution of pdays") +   theme_economist() + geom_density(alpha=.3) + scale_fill_discrete(name = "Attrition:")


ggplot(data=BankData,aes(x=BankData$previous, fill=BankData$y)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$previous, na.rm = TRUE), sd = sd(BankData$campaign, na.rm = TRUE))) +
  labs(x = "previous", y = "Count", title = "Distribution of previous") +   theme_economist() + geom_density(alpha=.3) + scale_fill_discrete(name = "Attrition:")


ggplot(data=BankData,aes(x=BankData$emp.var.rate, fill=BankData$y)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$emp.var.rate, na.rm = TRUE), sd = sd(BankData$emp.var.rate, na.rm = TRUE))) +
  labs(x = "emp.var.rate", y = "Count", title = "Distribution of emp.var.rate") +   theme_economist() + geom_density(alpha=.3) + scale_fill_discrete(name = "Attrition:")

ggplot(data=BankData,aes(x=BankData$cons.price.idx, fill=BankData$y)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$emp.var.rate, na.rm = TRUE), sd = sd(BankData$cons.price.idx, na.rm = TRUE))) +
  labs(x = "cons.price.idx", y = "Count", title = "Distribution of cons.price.idx") +   theme_economist() + geom_density(alpha=.3) + scale_fill_discrete(name = "Attrition:")

ggplot(data=BankData,aes(x=BankData$cons.conf.idx, fill=BankData$y)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$cons.conf.idx, na.rm = TRUE), sd = sd(BankData$cons.conf.idx, na.rm = TRUE))) +
  labs(x = "cons.conf.idx", y = "Count", title = "Distribution of cons.conf.idx") +   theme_economist() + geom_density(alpha=.3) + scale_fill_discrete(name = "Attrition:")

ggplot(data=BankData,aes(x=BankData$euribor3m, fill=BankData$y)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$euribor3m, na.rm = TRUE), sd = sd(BankData$euribor3m, na.rm = TRUE))) +
  labs(x = "euribor3m", y = "Count", title = "Distribution of euribor3m") +   theme_economist() + geom_density(alpha=.3) + scale_fill_discrete(name = "Attrition:")

ggplot(data=BankData,aes(x=BankData$nr.employed, fill=BankData$y)) + geom_histogram() +
  stat_function(fun = dnorm, args = list(mean = mean(BankData$nr.employed, na.rm = TRUE), sd = sd(BankData$nr.employed, na.rm = TRUE))) +
  labs(x = "nr.employed", y = "Count", title = "Distribution of nr.employed") +   theme_economist() + geom_density(alpha=.3) + scale_fill_discrete(name = "Attrition:")


```
### comments on histograms with the response added in
duration = close to zero seems like a really good indicator of failure.
pdays = 0 looks like a really good indication of success
previousl  = 4 also looks like a good indicator of success? Also maybe 2 and 3. Can tell about 5.
emp.var.rate being -3 or less looks like a good indicator of success.
cons.price.idx < 93 or so looks like a good indicator of success. It's odd that there are some values with very little no's and some with ton's of no's. Why is that? E.g. exactly 93 is like 50/50 yes and no, but a little less or more than 93 and it's a lot more no's. Why?
cons.conf.idx - again there are some values with very few no's - -50, -35 and -30. Then a lot of no's around -47, -44, -37. Why????
euribor3m - not sure whether this is useful. Definitely lower # gives higher probability of success

One thing to keep in mind with this data, especially this economic data, is that it represents ups and downs of the economy. But we are just seeing it FLAT, as "when it was at this, this happeneed". But it's really a sequence of events... E.g. if the rate is 5 and it goes to 4 it might generate more loans than it's it's 3 and it goes to 4?

### curious about comparing duration (how long was last contact) vs contact (cell phone vs. ...)

```{r}

ggplot(data=BankData, aes(BankData$contact, fill=BankData$y)) + 
  geom_bar(aes(y = (..count..)/sum(..count..)), position="dodge") + ylab("percent") + theme_economist() + ggtitle("Histogram of Marital Status") + xlab("Marital Status") + scale_fill_discrete(name = "Output:")

BankData %>% filter(contact == 'cellular')


BankData %>% filter(contact == 'cellular') %>%
ggplot(aes(x=duration, fill=y)) + geom_histogram() +
  labs(x = "duration", y = "Count", title = "Distribution of duration for cellular") +   theme_economist() + geom_density(alpha=.3) + scale_fill_discrete(name = "Attrition:")

BankData %>% filter(contact == 'telephone') %>%
ggplot(aes(x=duration, fill=y)) + geom_histogram() +
  labs(x = "duration", y = "Count", title = "Distribution of duration for telephone") +   theme_economist() + geom_density(alpha=.3) + scale_fill_discrete(name = "Attrition:")


```





## Some interactions between the data
  
```{r}
ggplot(BankData, aes(x=BankData$pdays, y=BankData$previous, color=BankData$y)) + geom_point() + geom_smooth(method=lm) +
  xlab("pdays - number of days since client contacted by previous campaign") + ylab("previous - number of contacts before this campaign") + ggtitle("pdays vs previous")

ggplot(BankData, aes(x=BankData$duration, y=BankData$campaign, color=BankData$y)) + geom_point() + geom_smooth(method=lm) +
  xlab("duration - last contact duration") + ylab("campaign - number of contacts during this campaign ") + ggtitle("duration vs campaign")

# economic data

ggplot(BankData, aes(x=BankData$emp.var.rate, y=BankData$nr.employed, color=BankData$y)) + geom_point() + geom_smooth(method=lm) +
  xlab("emp.var.rate") + ylab("nr.employed") + geom_jitter() + ggtitle("emp.var.rate vs nr.employed")

ggplot(BankData, aes(x=BankData$cons.price.idx, y=BankData$cons.conf.idx, color=BankData$y)) + geom_point() + geom_smooth(method=lm) +
  xlab("cons.price.idx") + ylab("cons.conf.idx") + geom_jitter() + ggtitle("cons.price.idx vs cons.conf.idx")

ggplot(BankData, aes(x=BankData$cons.price.idx, y=BankData$cons.conf.idx, color=BankData$y)) + geom_point() + geom_smooth(method=loess) +
  xlab("cons.price.idx") + ylab("cons.conf.idx") + geom_jitter() + ggtitle("cons.price.idx vs cons.conf.idx")

ggplot(BankData, aes(x=BankData$euribor3m, y=BankData$cons.conf.idx, color=BankData$y)) + geom_point() + geom_smooth(method=lm) +
  xlab("euribor3m") + ylab("cons.conf.idx") + geom_jitter() + ggtitle("euribor3m vs cons.conf.idx")

ggplot(BankData, aes(x=BankData$euribor3m, y=BankData$cons.conf.idx, color=BankData$y)) + geom_point() + geom_smooth(method=loess) +
  xlab("euribor3m") + ylab("cons.conf.idx") + geom_jitter() + ggtitle("euribor3m vs cons.conf.idx")

ggplot(BankData, aes(x=BankData$euribor3m, y=BankData$cons.price.idx, color=BankData$y)) + geom_point() + geom_smooth(method=lm) +
  xlab("euribor3m") + ylab("cons.price.idx") + geom_jitter() + ggtitle("euribor3m vs cons.price.idx")

ggplot(BankData, aes(x=BankData$euribor3m, y=BankData$cons.price.idx, color=BankData$y)) + geom_point() + geom_smooth(method=loess) +
  xlab("euribor3m") + ylab("cons.price.idx") + geom_jitter() + ggtitle("euribor3m vs cons.price.idx")


```

### comments on the plots
THe duration vs campaign is interesting. It shows that with the successful results, as the # of contacts increases, so did the duration of the last contact. But this is only the FINAL contact. So this might just be indicating that the when the final contact was LONG it is because it was success (e.g. them talking about the contract)? Is that final long duration contact really a good PREDICTOR then, since it could be occuring just when they go to "yes"? E.g. maybe it takes a while to do some paperwork or something if they say "yes"...


# run a basic logistic regression

```{r}

head(BankData)
head(BankData[c(21)])

# Function that returns Root Mean Squared Error
rmse <- function(error)
{
    sqrt(mean(error^2))
}

# remove the variable "Duration" because it is not available until after the data is collected, and is a giveaway (column 11)
includeForAuto = c(-11)
#includeForAuto = c(1,21)

#BankData[includeForAuto]


suppressMessages(library(MASS))
# Fit the full model 
logRegModel <- glm(y ~ ., data=BankData[includeForAuto], family = "binomial") %>%
  stepAIC(trace = FALSE)

summary(logRegModel)

# Stepwise regression model
step.model <- stepAIC(full.model, direction = "forward", trace = FALSE)
summary(step.model)

predictAttrLogisticRegresionProbs = predict(logRegModel, BankData, inverval="predict", type = "response")

predictAttrLogisticRegresion <- ifelse(predictAttrLogisticRegresionProbs > 0.5, "Yes", "No")


predictAttrLogisticRegresion <- as.factor(predictAttrLogisticRegresion)
length(predictAttrLogisticRegresion)


CM = confusionMatrix(table(predictAttrLogisticRegresion,trainAndMonthly$y))
CM

# let's lot the ROC curve

ROCRPred <- prediction(predictAttrLogisticRegresionProbs, trainAndMonthly$y)
ROCRPerf <- performance(ROCRPred,"tpr","fpr")
plot(ROCRPerf,colorize = TRUE, print.cutoffs.at = seq(0.1, by = 0.1))


```



```{r}
library("ROCR")
library("caret")

suppressMessages(library(ROCR))

# use the value with the monthly income filled in! Use our calculated values - they may be useful for predicting attrition
trainAndMonthly = rbind(train, predictMonthlyIncomeFilledIn)

includeForLogReg = c(2,3,4,5,6,7,8,9,12,13,14,15,16,17,18,19,20,21,22,24,25,26,27,28,29.30,31,32,33,34,35,36,37,38,39)

logRegModel <- glm(Attrition ~ ., data=trainAndMonthly[includeForLogReg], family = "binomial") %>%
  stepAIC(trace = FALSE)
summary(logRegModel)

predictAttrLogisticRegresionProbs = predict(logRegModel, trainAndMonthly, inverval="predict", type = "response")

predictAttrLogisticRegresion <- ifelse(predictAttrLogisticRegresionProbs > 0.2, "yes", "no")


predictAttrLogisticRegresion <- as.factor(predictAttrLogisticRegresion)
length(predictAttrLogisticRegresion)


CM = confusionMatrix(table(predictAttrLogisticRegresion,BankData$y))
CM

# let's lot the ROC curve

ROCRPred <- prediction(predictAttrLogisticRegresionProbs, trainAndMonthly$Attrition)
ROCRPerf <- performance(ROCRPred,"tpr","fpr")
plot(ROCRPerf,colorize = TRUE, print.cutoffs.at = seq(0.1, by = 0.1))


```



